---
title: "'JSON' columns to data frame"
subtitle: "For R users"
author: "Georgi Petkov"
date: "04 March 2019"
output:
 html_document:
    toc: true
    theme: cosmo
    highlight: tango
    code_folding: hide

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview

TMDB Box Office Prediction competition comes with data in a wired format, at least to me.
Maybe, I am missing something but some of the columns are in a format that is supposed to be JSON,
I believe so because other similar datasets have it this way, but the single quotation will not let us use
"jsonlight" package to convert it to a data frame. The following code is my solution to the problem.
If you find it useful please upvote or if you have any suggestions leave a comment.

## Load libraries
```{r,echo=TRUE,message=FALSE,warning=FALSE}
library(tidyverse)
library(DT)
```

## Load data
```{r, echo=FALSE,message=FALSE}
train <- read_csv("../input/train.csv")

train %>%
  glimpse()
```

## Preprocessing and sneak peek of the data
The way I've decided to go about this problem is to extract the data, split the string, name each element and put everything in a df
for each "JSON" column.

## Cast

```{r,echo=FALSE,message=FALSE,warning=FALSE}
cast_list <- list()

for (i in seq_along(train$cast)) {
  cast_list[[i]] <- train$cast[[i]] %>%
    str_extract_all('(?<=\\{).*?(?=\\})') %>%  #extract everything between {}
    str_split("(?<=[:digit:]|[:punct:]), ",
              n=Inf,simplify = TRUE) %>%       #split on ","
    str_extract_all('(?<=\\:[:space:]).*') %>% #get the part after the semicolon
    str_replace_all("'|\"","") %>% #clean the unwanted punctuation
    matrix( ncol = 8,  byrow = TRUE,dimnames=list(c(),
                                                  c("cast_id","character","credit_id","gender","id",
                                                    "name","order","profile_path"))) %>% #convert to matrix
    as_tibble(stringsAsFactors = FALSE)#convert the matrix to tibble
}

# Name the list
names(cast_list) <- c(1:3000) #name the list with the sequential number of the list element

# Create df with a column that identifies each movie
cast_df <- bind_rows(cast_list, .id = 'movie_id')#create single df with a column that identifies the movies


cast_df %>% 
  head(10) %>% 
  DT::datatable()

```

## Genres

```{r,echo=FALSE,message=FALSE,warning=FALSE}
genres_list <- list()
for (i in seq_along(train$genres)) {
  genres_list[[i]] <- train$genres[[i]] %>%
    str_extract_all('(?<=\\{).*?(?=\\})') %>% 
    str_split("(?<=[:digit:]|[:punct:]), ",n=Inf,simplify = TRUE) %>% 
    str_extract_all('(?<=\\:[:space:]).*') %>% 
    str_replace_all("[:punct:]","") %>% 
    matrix( ncol = 2,  byrow = TRUE,dimnames=list(c(),
                                                  c("id","name"))) %>% 
    as_tibble(stringsAsFactors = FALSE)
}

names(genres_list) <- c(1:3000)

genres_df <- bind_rows(genres_list, .id = 'movie_id')

genres_df %>% 
  head(10) %>% 
  DT::datatable()
```

## Production companies

```{r,echo=FALSE,message=FALSE,warning=FALSE}
production_companies_list <- list()
for (i in seq_along(train$production_companies)) {
  production_companies_list[[i]] <- train$production_companies[[i]] %>%
    str_extract_all('(?<=\\{).*?(?=\\})') %>% 
    str_split("(?<=[:digit:]|[:punct:]), ",n=Inf,simplify = TRUE) %>% 
    str_extract_all('(?<=\\:[:space:]).*') %>% 
    str_replace_all("[:punct:]","") %>% 
    matrix( ncol = 2,  byrow = TRUE,dimnames=list(c(),
                                                  c("name","id"))) %>% 
    as_tibble(stringsAsFactors = FALSE)
}

names(production_companies_list) <- c(1:3000)

production_companies_df <- bind_rows(production_companies_list, .id = 'movie_id')


production_companies_df %>% 
  head(10) %>% 
  DT::datatable()
```


## Crew
```{r,echo=FALSE,message=FALSE,warning=FALSE}

crew_list <- list()

for (i in seq_along(train$crew)) {
  crew_list[[i]] <- train$crew[[i]] %>%
    str_extract_all('(?<=\\{).*?(?=\\})') %>% 
    str_split("(?<=[:digit:]|[:punct:]), ",n=Inf,simplify = TRUE) %>% 
    str_extract_all('(?<=\\:[:space:]).*') %>% 
    str_replace_all("[:punct:]","") %>% 
    matrix( ncol = 7,  byrow = TRUE,dimnames=list(c(),
   c("credit_id","department","gender","id","job","name","profile_path"))) %>% 
    as_tibble(stringsAsFactors = FALSE)
}

names(crew_list) <- c(1:3000)

crew_df <- bind_rows(crew_list, .id = 'movie_id')

crew_df%>% 
  head(10) %>% 
  DT::datatable()
```

I will assume that if anybody had a problem with figuring out how to extract the data from these columns this kernel will make his life easier.

